%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plasmati Graduate CV
% LaTeX Template
% Version 1.0 (24/3/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Alessandro Plasmati (alessandro.plasmati@gmail.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
% Important note:
% This template needs to be compiled with XeLaTeX.
% The main document font is called Fontin and can be downloaded for free
% from here: http://www.exljbris.com/fontin.html
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
% Default font size and paper size
\documentclass[letterpaper,12pt]{article} 

% For loading fonts
\usepackage{fontspec} 
\defaultfontfeatures{Mapping=tex-text}
\setmainfont[SmallCapsFont = Fontin-SmallCaps.ttf, BoldFont=Fontin-Bold.ttf, ItalicFont=Fontin-Italic.ttf]{Fontin-Regular.ttf} % Main document font

% Formatting packages
\usepackage{xunicode,xltxtra,url,parskip} 

% Required for specifying custom colors
\usepackage[usenames,dvipsnames]{xcolor}

% Required for adding links	and customizing them
\usepackage{hyperref}
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
% Set link colors throughout the document
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour}

% Used to customize the \section command
\usepackage{titlesec}
% Text formatting of sections
\titleformat{\section}{\Large\scshape\raggedright}{}{0em}{}[\titlerule]
% Spacing around sections
\titlespacing{\section}{0pt}{3pt}{3pt}

\usepackage{array}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage[margin=0.5in]{geometry}
\usepackage{setspace}

\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

\begin{document}

% Removes page numbering
\pagestyle{empty}

%----------------------------------------------------------------------------------------
%	NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\par{\centering{\Huge Dennis F. \textsc{Gardner Jr.}, Ph.D.}\bigskip\par}

\vspace{-0.5cm}
\begin{table}[h!]
	\centering
	\begin{tabular}{c | c | c | c }
		Vienna, VA & 970-412-4875 & \href{mailto:dennisfgardner@gmail.com}{dennisfgardner@gmail.com} & \href{https://scholar.google.com/citations?user=uYXouIIAAAAJ}{Publications \& Patents}
	\end{tabular}
\end{table}

\section{Professional Summary}
\vspace{0.2cm}

Physicist and systems engineer with 10+ years advancing sensing, imaging, and autonomous systems for national security missions.
Equally effective in the lab and in the boardroom---from hands-on experimentation and systems architecture to program leadership and executive briefings.
Recognized for tackling ambiguous, cross-disciplinary hardware/software problems and translating complex technical work into mission impact and business growth.

\section{Areas of Expertise}
\begin{table}[h!]
	\centering
	\begin{tabular}{l | l  }
		\checkmark Autonomous Systems & \checkmark Artificial Intelligence \& Machine Learning \\[0.2cm]
		\checkmark Highly Performant Code  & \checkmark Cloud Computing (AWS EC2, S3, SQS) \\[0.2cm]
		\checkmark Data Analysis \& Visualization & \checkmark Computer Vision (OpenCV)  \\[0.2cm]
		\checkmark Software Design \& Engineering & \checkmark Computer/Device Interfacing (ROS2) \\[0.2cm]
		\checkmark Presentations \& Public Speaking & \checkmark Scientific \& Technical Writing  \\
	\end{tabular}
\end{table}

\section{Work Experience}
\begin{tabular}{rp{13cm}}
    \textsc{May 2025}--\textsc{May 2026} & Principal Scientific Director at \textsc{Ancora Collective}  \\  
    & \emph{Scientific Consulting} \\   
    & \footnotesize{ 
        \begin{itemize}[topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt]
            \vspace{-0.2cm}
            \item Led systems engineering and analysis for complex, multi-disciplinary technical programs, translating stakeholder requirements into system architectures and design solutions
            \item Served as technical lead and subject-matter expert interfacing between scientific/engineering teams and non-technical stakeholders
            \item Advised on system design trade-offs and technology strategy, integrating insights from emerging technologies into national security decision-making
    \end{itemize}}\\
\end{tabular}

\begin{tabular}{rp{13cm}}
    \textsc{Apr 2024}--\textsc{May 2025} & Solutions Lead Senior Principal at \textsc{MITRE}  \\  
    & \emph{Systems Engineering} \\   
    & \footnotesize{ 
        \begin{itemize}[topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt]
            \vspace{-0.2cm}
			\item Secured multi-million-dollar ONR funding by identifying a sponsor capability gap, shaping the technical solution, and leading the project through MITRE's innovation pipeline from concept validation to funded program
            \item Pioneered a trajectory-prediction capability using LLM architectures, leading development from novel research concept to a live demonstration for a government sponsor
            \item Generated two patent applications from portfolio research, translating emerging technology into protectable intellectual property for national security applications
    \end{itemize}}\\
\end{tabular}

\begin{tabular}{rp{13cm}}
	\textsc{Mar 2023}--\textsc{Apr 2024} & Chief Scientist of EO Remote Sensing and Imaging at \textsc{GDIT}  \\  
	& \emph{Machine Learning \& Computer Vision ATR} \\   
	& \footnotesize{ 
		\begin{itemize}[topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt]
			\vspace{-0.2cm}
			\item Served as the technical authority on the development of automated ATR and anomaly-detection algorithms for large geospatial imagery datasets
			\item Transitioned prototype computer-vision algorithms into the production code base, defining performance metrics and evaluation processes used to validate capability under operational conditions
			\item Briefed customers on AI/ML integration with EO/IR sensing and supported technical proposals that expanded expanded contract scope
	\end{itemize}}\\
\end{tabular}

\begin{tabular}{rp{13cm}}
	\textsc{Jan 2020}--\textsc{Mar 2023} & Lead Computational Imaging Scientist at \textsc{MITRE}  \\  
	& \emph{Rapid Prototyping} \\   
	& \footnotesize{ 
	\begin{itemize}[topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt]
		\vspace{-0.2cm}
		\item Through-wall radar sensing and tracking
		\item LIDAR-based obstacle avoidance for autonomous mine-clearing vehicles
		\item Non-contact biometric collection
	\end{itemize}}\\
\end{tabular}

\begin{tabular}{rp{13cm}}
\textsc{Feb 2017}--\textsc{Jan 2020} & Research Physicist at \textsc{U.S. Naval Research Laboratory}  \\  
& \emph{Computational Optics, Sensing, and Imaging} \\   
& \footnotesize{ 
\begin{itemize}[topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt]
	\vspace{-0.2cm}
	\item Adaptive Optics (AO) simulation in extended volume turbulence
	\item Synthetic Aperture Radar (SAR) image analysis and anomaly detection onboard autonomous systems
	% \item Orbital Angular Momentum (OAM) beams for underwater communications
	\item Structured illumination for night vision applications
	\item Non-Line of Sight (NLOS) imaging using speckle correlations
	\item Holographic imaging through fog and turbid water
\end{itemize}}\\
\end{tabular}

\begin{tabular}{rp{13cm}}
\textsc{Aug 2010}--\textsc{Dec 2016} & Research Assistant at \textsc{University of Colorado at Boulder}\\
& \emph{Lensless High-Resolution Imaging with Extreme Ultraviolet Light}\\ 
& \footnotesize{
	\begin{itemize}[topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt]
		\vspace{-0.2cm}
		\item Extreme Ultraviolet (EUV) light generation using the high-harmonic process
		\item High-resolution Coherent Diffraction Imaging (CDI)
		\item Pump-probe stroboscopic experiments 

\end{itemize}}\\
\end{tabular}

\begin{tabular}{rp{13cm}}
    \textsc{Jun 2012}--\textsc{Aug 2012} & Internship at \textsc{Lawrence Livermore National Lab}  \\  
    & \emph{Department of Homeland Security Summer Program} \\   
    & \footnotesize{ 
        \begin{itemize}[topsep=0pt,itemsep=0pt,partopsep=0pt,parsep=0pt]
            \vspace{-0.2cm}
            \item Hardware assembly for Free-Electron Laser (FEL) experiments
            \item Data analysis and visualization
    \end{itemize}}\\
\end{tabular}

\vspace{-0.1cm}
\section{Awards, Fellowships \& Scholarships}
\begin{table}[h!]
	\centering
	\begin{tabular}{l | l  }
		 2009 Mensa Scholarship &  2010 SPIE Scholarship\\
		 2011 NSF COSI Traineeship &  2011 Ford Foundation Fellowship \footnotesize(2-year fellowship)\normalsize \\
		 2011 NSF GRFP \footnotesize(3-year fellowship)\normalsize & 2015 OSA Foundation's Outstanding Student Paper\\
		 2017 APS Dissertation Prize in Laser Science &  2017 NRL Postdoc Fellowship \footnotesize(2-year fellowship)\normalsize\\
		 2018 NRL Research Publication Award &  2022 MITRE Trailblazer Award  \\
		 2024 MITRE Breakthrough Award & \\
	\end{tabular}
\end{table}

\vspace{-0.1cm}
\section{Education}
% \vspace{0.2cm}
\begin{tabular}{rl}	
	\textsc{Dec} 2016 & Doctor of Philosophy in \textbf{Physics}, \textsc{The University of Colorado at Boulder}\\
	\vspace{0.5cm}
	& \small\emph{Received dissertation award from the American Physical Society }\\
	\textsc{May} 2010 & Bachelor of Arts in \textbf{Physics}, \textsc{The University of Colorado at Boulder}\\
	& \small \textit{Graduated with summa cum laude honors}\\
\end{tabular}

\vspace{-0.3cm}
\section{Patents}
\vspace{0.2cm}
\begin{enumerate}[leftmargin=*]
	\item A Watnik, \textbf{DF Gardner}, T Dubose, ``Methods and apparatuses for wavefront reconstruction,'' \emph{US Patent 12,432,452, 2025}.
    \item R Houvener, G Cardinale, R Lamond, M Stenner, D Moriarty, C Zhao, \& \textbf{DF Gardner}, ``Scalable Coherent Laser Array Systems,'' \emph{US Patent App. 18/927,444, 2024}.
	\item  M Tanksalvala, DE Adams, \textbf{DF Gardner}, CL Porter, \& GF Mancini, ``Modulus-Enforced Probe,'' \emph{US Patent 11,709,132, 2023}.
	\item S Durvasula, \textbf{DF Gardner}, \emph{et al.}, ``Systems and methods for remote optical screening of individuals in entryways,'' \emph{US Patent App. 18/111,312, 2023}.
	\item \textbf{DF Gardner}, ``Apparatuses and Methods for Imaging Incoherently Illuminated Objects,'' \emph{US Patent App. 17/331,932, 2021}.
\end{enumerate}

\section{Publications -- 19 peer-reviewed, 1,900+ citations, h-index 18}
\vspace{0.2cm}

% \textbf{19 peer-reviewed publications, 1,900+ citations, h-index of 18}

\begin{enumerate}[leftmargin=*]
	
	\item S Divitt, S Park, H Gemar, KP Judd, H Dave, D Lin, \textbf{DF Gardner}, AT Watnik, ``Structured illumination and image enhancement of three-dimensional and moving objects at a distance via incoherent Fourier Ptychography,'' \emph{Applied Optics} \textbf{63}(8), C8-C14 (2024). \href{https://doi.org/10.1364/AO.505275}{link}
	
	\item TB DuBose, \textbf{DF Gardner}, AT Watnik, ``Intensity-enhanced deep network wavefront reconstruction in Shack–Hartmann sensors,'' \emph{Optics Letters} \textbf{45}(7), 1699-1702 (2020). \href{https://doi.org/10.1364/OL.389895}{link}

	\item S Divitt, \textbf{DF Gardner}, AT Watnik, ``Imaging around corners in the mid-infrared using speckle correlations,'' \emph{Optics Express} \textbf{28}(8), 11051-11064 (2020). \href{https://doi.org/10.1364/OE.388260}{link}

	\item \textbf{DF Gardner}, S Divitt, AT Watnik, ``Ptychographic imaging of incoherently illuminated extended objects using speckle correlations,'' \emph{Applied Optics} \textbf{58}(13), 3564-3569 (2019). \href{https://doi.org/10.1364/AO.58.003564}{link}
	
	\item AV Kanaev, AT Watnik, \textbf{DF Gardner}, C Metzler, KP Judd, P Lebow, KM Novak, \& JR Lindle, ``Imaging through extreme scattering in extended dynamic media,''  \emph{Optics Letters} \textbf{43}(13), 3088-3091 (2018). \href{https://doi.org/10.1364/OL.43.003088}{link}
	
	\quad $\rightarrow$ \textbf{Paper highlighted in OSA's 2018 Year in Review and winner of 2018 NRL publication award}
	
	\item RM Karl, GF Mancini, JL Knobloch, TD Frazer, JN Hernandez-Charpak, B Abad, \textbf{DF Gardner}, ER Shanblatt, M Tanksalvala, CL Porter, CS Bevis, DE Adams, HC Kapteyn, \& MM Murnane, ``Full-field imaging of thermal and acoustic dynamics in an individual nanostructure using tabletop high harmonic beams,'' \emph{Science Advances} \textbf{4}(10), eaau4295 (2018). \href{https://doi.org/10.1126/sciadv.aau4295}{link}
	
	\item  C Bevis, R Karl Jr., J Reichanadter, \textbf{DF Gardner}, C Porter, E Shanblatt, M Tanksalvala, GF Mancini, H Kapteyn, M Murnane \& D Adams, ``Multiple beam ptychography for large field-of-view, high throughput, quantitative phase contrast imaging,''  \emph{Ultramicroscopy} \textbf{184}(A), 164-171 (2018). \href{https://doi.org/10.1016/j.ultramic.2017.08.018}{link}
	
	\item A Pryor, A Rana, R Xu, JA Rodriguez, Y Yang, M Gallagher-Jones, H Jiang, K Kanhaiya, M Nathanson, J Park, S Kim, S Kim, D Nam, Y Yue, J Fan, Z Sun, B Zhang, \textbf{DF Gardner}, CSB Dias, Y Joti, T Hatsui, T Kameshima, Y Inubushi, K Tono, JY Lee, M Yabashi, C Song, T Ishikawa, HC Kapteyn, MM Murnane, H Heinz, J Miao, ``Single-shot 3D coherent diffractive imaging of core-shell nanoparticles with elemental specificity,'' \emph{Scientific Reports} \textbf{8}(1), 8284 (2018). \href{https://doi.org/10.1038/s41598-018-26182-1}{link}
	
	\quad $\rightarrow$ \textbf{Collaboration with UCLA and the SACLA free-electron laser facility in Japan}
	
	\item \textbf{DF Gardner}, M Tanksalvala, ER Shanblatt, X Zhang, BR Galloway, CL Porter, R Karl Jr, C Bevis, DE Adams, HC Kapteyn, MM Murnane \& GF Mancini, ``Subwavelength coherent imaging of periodic samples using a 13.5 nm tabletop high harmonic light source,'' \emph{Nature Photonics} \textbf{11}, 259-263 (2017). \href{https://doi.org/10.1038/nphoton.2017.33}{link}
	
	\quad $\rightarrow$ \textbf{Paper highlighted in OSA's 2017 Year in Review}
	
	\quad $\rightarrow$ \textbf{This paper has over 150 citations}
	
	\item ER Shanblatt, CL Porter, \textbf{DF Gardner}, GF Mancini, RM Karl Jr, MD Tanksalvala, CS Bevis, VH Vartanian, HC Kapteyn, DE Adams \& MM Murnane, ``Quantitative chemically specific coherent diffractive imaging of reactions at buried interfaces with few nanometer precision,'' \emph{Nano Letters} \textbf{16}(9), 5444-5450 (2016). \href{https://doi.org/10.1021/acs.nanolett.6b01864}{link}
	
	\item B Zhang, \textbf{DF Gardner}, MH Seaberg, ER Shanblatt, CL Porter, R Karl, CA Mancuso, HC Kapteyn, MM Murnane \& DE Adams, ``Ptychographic hyperspectral spectromicroscopy with an extreme ultraviolet high harmonic comb,''  \emph{Optics Express} \textbf{24}(16), 18745-18754 (2016). \href{https://doi.org/10.1364/OE.24.018745}{link}
	
	\item R Karl, C Bevis, R Lopez-Rios, J Reichanadter, \textbf{DF Gardner}, C Porter, E Shanblatt, M Tanksalvala, GF Mancini, M Murnane, H Kapteyn \& D Adams, ``Spatial, spectral, and polarization multiplexed ptychography,'' \emph{Optics Express} \textbf{23}(23), 30250-30258 (2015). \href{https://doi.org/10.1364/OE.23.030250}{link}
	
	\item B Zhang, \textbf{DF Gardner}, MD Seaberg, ER Shanblatt, HC Kapteyn, MM Murnane \& DE Adams, ``High contrast 3D imaging of surfaces near the wavelength limit using tabletop EUV ptychography,'' \emph{Ultramicroscopy} \textbf{158}, 98-104 (2015). \href{https://doi.org/10.1016/j.ultramic.2015.07.006}{link}
	
	\quad $\rightarrow$ \textbf{Work featured on journal and Laser Focus World covers}
	
	\quad $\rightarrow$ \textbf{This paper has over 75 citations}
	
	\item  MD Seaberg, B Zhang, \textbf{DF Gardner}, ER Shanblatt, MM Murnane, HC Kapteyn \& DE Adams, ``Tabletop nanometer extreme ultraviolet imaging in an extended reflection mode using coherent Fresnel ptychography,''  \emph{Optica} \textbf{1}(1), 39-44 (2014). \href{https://doi.org/10.1364/OPTICA.1.000039}{link}
	
	\quad $\rightarrow$ \textbf{This paper has over 125 citations}
	
	\item T Pardini, S Boutet, J Bradley, T Döppner, LB Fletcher, \textbf{DF Gardner}, RM Hill, MS Hunter, J Krzywinski, M Messerschmidt, AE Pak, F Quirin, K Sokolowski-Tinten, GJ Williams, SP Hau-Riege, ``Silicon mirrors for high-intensity x-ray pump and probe experiments,''  \emph{Physical Review Applied} \textbf{1}(4), 044007 (2014). \href{https://doi.org/10.1103/PhysRevApplied.1.044007}{link}
	
	\quad $\rightarrow$ \textbf{Collaboration with LLNL and LCLS free-electron laser facility at SLAC}
	
	\item B Zhang, MD Seaberg, DE Adams, \textbf{DF Gardner}, ER Shanblatt, JM Shaw, W Chao, EM Gullikson, F Salmassi, HC Kapteyn \& MM Murnane, ``Full field tabletop EUV coherent diffractive imaging in a transmission geometry,'' \emph{Optics Express} \textbf{21}(19), 21970-21980 (2013). \href{https://doi.org/10.1364/OE.21.021970}{link}
	
	\item  DE Adams, LS Martin, MD Seaberg, \textbf{DF Gardner}, HC Kapteyn \& MM Murnane, ``A generalization for optimized phase retrieval algorithms,'' \emph{Optics Express} \textbf{20}(22), 24778-24790 (2012). \href{https://doi.org/10.1364/OE.20.024778}{link}
	
	\item \textbf{DF Gardner}, B Zhang, MD Seaberg, LS Martin, DE Adams, F Salmassi, E Gullikson, HC Kapteyn \& MM Murnane, ``High numerical aperture reflection mode coherent diffraction microscopy using off-axis apertured illumination,'' \emph{Optics Express} \textbf{20}(17), 19050-19059 (2012). \href{https://doi.org/10.1364/OE.20.019050}{link}
	
	\quad $\rightarrow$ \textbf{This paper has over 75 citations}
	
	\item  Q Liu, Y Cui, \textbf{DF Gardner}, X Li, S He \& II Smalyukh, ``Self-alignment of plasmonic gold nanorods in reconfigurable anisotropic fluids for tunable bulk metamaterial applications,''  \emph{Nano Letters} \textbf{10}(4), 1347-1353 (2010). \href{https://doi.org/10.1021/nl9042104}{link}
	
	\quad $\rightarrow$ \textbf{This paper has over 350 citations}

\end{enumerate}

\section{Other Publications}
\begin{enumerate}[leftmargin=*]
	
	\item AT Watnik \& \textbf{DF Gardner}, ``Wavefront Sensing in Deep Turbulence,''  \emph{Optics and Photonics News} \textbf{29}(10), 39-45 (2018).
	
	\quad $\rightarrow$ \textbf{Article featured on the cover}
	
	\item AV Kanaev, AT Watnik, \textbf{DF Gardner}, KP Judd, P Lebow, JR Lindle, C Metzler, \& KM Novak, ``Imaging Through Macroscopic Scattering Media,'' \emph{Optics and Photonics News} \textbf{29}(12), 30-31 (2018).
	
	\item M Tanksalvala, C Porter, \textbf{DF Gardner}, X Zhang, E Shanblatt, B Galloway, R Karl Jr., C Bevis, G Mancini, D Adams, H Kapteyn \& M Murnane, ``Subwavelength EUV Imaging,'' \emph{Optics and Photonics News} \textbf{28}(12), 28-29 (2017).
	
\end{enumerate}

\end{document}
